Figures

### 2016-08-10 coralmaps/coral_maps.Rmd
fixing up the plots: new quad plot, new coral quad plot, new dataset by taxa plot, new OHI plot

if(!file.exists(file.path(dir_git, 'coralmaps', 'coral_spp_depth_raw.csv'))) {

@@

@@

spp_coralmaps <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_areas.csv'))

### check raw iucn_area vs clipped - looks proportional; scaling factor due to difference in area calcs?
scatterplot1 <- ggplot(spp_coralmaps, aes(x = area_iucn_raw, y = area_iucn_clipped)) +
  geom_point() +
  geom_abline(color = 'red')
scatterplot1

### rough recalcs
### assume that by removing the deep cells, we will not remove cells that are
### included in AquaMaps, since AquaMaps already filters by depth.  This means
### the overlap area will not be affected.
iucn_coral_fixed <- spp_coralmaps %>%
  gather(dist, perc, c(sm_perc_raw, sm_perc_clipped)) %>%
  gather(area, ratio, c(area_ratio_raw, area_ratio_clipped)) %>%
  filter(!(str_detect(dist, 'raw') & !(str_detect(area, 'raw')))) %>%
  filter(!(str_detect(dist, 'clip') & !(str_detect(area, 'clip')))) %>%
  mutate(method = ifelse(str_detect(dist, 'clip'), '200m clip', 'all depth'))

source(file.path(dir_git, 'data_explore/data_explore_fxns.R'))
clipped_quads <- ggplot(iucn_coral_fixed %>%
                      filter(!iucn_sid %in% (iucn_coral_fixed %>% 
                                               filter(perc > 100) %>% 
                                               .$iucn_sid)),
                    aes(x = ratio, y = perc, group = iucn_sid)) +
  ggtheme_plot + 
  ### color the quadrant backgrounds:
  # annotate("rect", xmin = 54.5, xmax = 100, ymin = 63, ymax = 100, 
  #          alpha = .3, fill= "#4dac26")  + 
  # annotate("rect", xmax = 54.5, xmin =   0, ymin = 63, ymax = 100, 
  #          alpha = .3, fill= "#b8e186") + 
  # annotate("rect", xmin = 54.5, xmax = 100, ymax = 63, ymin =   0, 
  #          alpha = .3, fill= "#f1b6da") + 
  # annotate("rect", xmax = 54.5, xmin =   0, ymax = 63, ymin =   0, 
  #          alpha = .3, fill= "#d01c8b") +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 102),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%')) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 103),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%')) +

  geom_line(color = 'grey50', size = .5, alpha = .3) +
  geom_point(aes(color = method), size = 2, alpha = .8) +
  labs(x = 'Area ratio (smaller/larger)',
       y = 'Distribtion (%smaller in larger)')

>>>>

if(!file.exists(file.path(dir_git, 'coralmaps', 'coral_spp_depth.csv'))) {

@@

} else 
  spp_depth <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_depth.csv'))

@@

spp_coralmaps <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_areas.csv')) %>% 
                      mutate(loss_ratio = 100 * area_iucn_clipped/area_iucn_raw)

spp_small_effect <- spp_coralmaps %>%
  filter(loss_ratio > 80) 
### only five with such small effects; one of those did not change at all.

summary(spp_coralmaps$loss_ratio)
  #  Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  # 23.80   52.89   55.48   56.33   60.59  100.00

hist(spp_coralmaps$loss_ratio)
### check raw iucn_area vs clipped - looks proportional; scaling factor due to difference in area calcs?

scatterplot1 <- ggplot(spp_coralmaps, aes(x = area_iucn_raw, y = area_iucn_clipped)) +
  geom_point() +
  geom_abline(color = 'red')
scatterplot1

violplot1 <- ggplot(spp_coralmaps, aes(x = 'sm_range_raw', y = loss_ratio)) +
  ggtheme_plot +
  geom_violin(fill = 'grey80', alpha = .5,
              draw_quantiles = c(.25, .5, .75)) +
  geom_point(position = position_jitter(w = .3), 
             color = '#4d4dac', alpha = .5) +
  geom_violin(fill = NA, draw_quantiles = c(.25, .5, .75)) +
  geom_hline(color = 'grey30', yintercept = 100) +
  scale_y_continuous(limits = c(0, 100)) +
  scale_x_discrete(labels = element_blank()) +
  theme(axis.title.x = element_blank()) +
  labs(y = 'IUCN coral range: (<= 200 m) / (total)')
print(scatterplot1)

### recalc areas and overlaps
iucn_coral_fixed <- spp_coralmaps %>%
  gather(dist, perc, c(sm_perc_raw, sm_perc_clipped)) %>%
  gather(area, ratio, c(area_ratio_raw, area_ratio_clipped)) %>%
  filter(!(str_detect(dist, 'raw') & !(str_detect(area, 'raw')))) %>%
  filter(!(str_detect(dist, 'clip') & !(str_detect(area, 'clip')))) %>%
  mutate(method = ifelse(str_detect(dist, 'clip'), '200 m clip', 'original'))

source(file.path(dir_git, 'data_explore/data_explore_fxns.R'))

break_nums <- seq(0, 100, 20)

clipped_quads <- ggplot(iucn_coral_fixed %>%
                      filter(!iucn_sid %in% (iucn_coral_fixed %>% 
                                               filter(perc > 100) %>% 
                                               .$iucn_sid)),
                    aes(x = ratio, y = perc, group = iucn_sid)) +
  ggtheme_plot + 
  ### color the quadrant backgrounds:
  annotate("rect", xmin = 54.5, xmax = 100, ymin = 63, ymax = 100,
           alpha = .3, fill= "#4dac26")  +
  annotate("rect", xmax = 54.5, xmin =   0, ymin = 63, ymax = 100,
           alpha = .3, fill= "#b8e186") +
  annotate("rect", xmin = 54.5, xmax = 100, ymax = 63, ymin =   0,
           alpha = .3, fill= "#f1b6da") +
  annotate("rect", xmax = 54.5, xmin =   0, ymax = 63, ymin =   0,
           alpha = .3, fill= "#d01c8b") +
  ### add grid lines
  geom_vline(xintercept = break_nums, size = 0.25, color = 'white', alpha = .5) +
  geom_hline(yintercept = break_nums, size = 0.25, color = 'white', alpha = .5) +

  scale_x_continuous(expand = c(0, 0), limits = c(-1, 101),
                     breaks = break_nums,
                     labels = sprintf('%s%%', break_nums)) +
  scale_y_continuous(expand = c(0, 0), limits = c(-1, 101),
                     breaks = break_nums,
                     labels = sprintf('%s%%', break_nums)) +

  geom_line(color = 'grey50', size = .25, alpha = .3) +
  geom_point(aes(color = method, size = method), alpha = 1) +
  scale_color_manual(values = c('original' = 'grey60', '200 m clip' = 'black')) +
  scale_size_manual(values = c('original' = .25, '200 m clip' = .25)) +
    labs(x = bquote('Area ratio '*alpha['ext']*' (%)'), 
         y = bquote('Distribution alignment '*alpha['dist']*' (%)'))

### 2016-08-10 data_explore/data_explore.Rmd
fixing up the plots: new quad plot, new coral quad plot, new dataset by taxa plot, new OHI plot

bar_taxa <- ggplot(taxa_df_long, aes(x = class, y = percent, fill = dataset)) +
  ggtheme_plot + 
  theme(panel.grid.major.x = element_blank()) +
  geom_bar(stat = 'identity', alpha = .9) +
  scale_fill_manual(values = colvec) +
  geom_text(aes(label = sprintf('n = %s', n_total), y = 102), 
            size = 2, 
            angle = 90, hjust = 0, color = 'grey30') +
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 130),
                     breaks = c(0, 25, 50, 75, 100),
                     labels = c('0%', '25%', '50%', '75%', '100%')) + 
  labs(# title = 'Species maps by taxonomic class',
       x = 'Class', 
       y = 'Percent of mapped species',
       fill = 'Dataset')

@@

@@

### mongo plot time
scatter_quadplot <- ggplot(spp_list_labs, aes(x = x, y = y)) +
  ggtheme_plot + 
  ### color the quadrant backgrounds:
  annotate("rect", xmin = area_align_mean, xmax = 100, ymin = dist_align_mean, ymax = 100, 
           alpha = .3, fill= "#4dac26")  + 
  annotate("rect", xmax = area_align_mean, xmin =   0, ymin = dist_align_mean, ymax = 100, 
           alpha = .3, fill= "#b8e186") + 
  annotate("rect", xmin = area_align_mean, xmax = 100, ymax = dist_align_mean, ymin =   0, 
           alpha = .3, fill= "#f1b6da") + 
  annotate("rect", xmax = area_align_mean, xmin =   0, ymax = dist_align_mean, ymin =   0, 
           alpha = .3, fill= "#d01c8b")

  ### plot the data points!
scatter_quadplot <- scatter_quadplot +
  # geom_point(color = 'white',   alpha = 1) + ### blank behind the points
  geom_point(color = '#4d4dac', aes(alpha = fade), show.legend = FALSE) +
  scale_alpha_manual(values = c('TRUE' = .15, 'FALSE' = .7))

@@

### Manage axis scales 
scatter_quadplot <- scatter_quadplot +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 102),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%')) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 103),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%'))

@@

### Plot the bar chart
barchart_spp_gp_quads <- ggplot(spp_gp_quadrants %>% 
                                  transform(spp_group_text = reorder(spp_group_text, pct_q1)), 
                                aes(x = spp_group_text, fill = quad, weight = pct_quad)) +
  ggtheme_plot + 
  theme(panel.grid.major.y = element_blank()) +
  geom_bar(stat = 'count', alpha = .8) +
  scale_fill_manual(values = c('q1'   = '#4dac26', 
                               'q2'   = '#b8e186', 
                               'q3'   = '#f1b6da', 
                               'q4' = '#d01c8b'),
                    labels = levels(quad_names)) +
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 1.1),
                     breaks = c(0, .25, .5, .75, 1.0),
                     labels = c('0%', '25%', '50%', '75%', '100%')) +
  geom_text(aes(label = sprintf('n = %s', n_spp), y = 1.01), hjust = 0, 
            size = 2, 
            color = 'grey30') +
  coord_flip() +
  labs(x = 'Taxonomic Group', 
       y = 'Percent of species by quadrant', 
       fill = 'Alignment')

@@

### Plot the bar chart
barchart_spp_risk_quads <- ggplot(spp_risk_quadrants1 %>% 
                                  transform(popn_category = reorder(popn_category, cat_rank)), 
                                aes(x = popn_category, fill = quad, weight = pct_quad)) +
  ggtheme_plot + 
  theme(panel.grid.major.y = element_blank()) +
  geom_bar(stat = 'count', alpha = .8) +
  scale_fill_manual(values = c('q1 well-matched'   = '#4dac26', 
                               'q2 dist-matched'   = '#b8e186', 
                               'q3 area-matched'   = '#f1b6da', 
                               'q4 poorly matched' = '#d01c8b')) +
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 1.1),
                     breaks = c(0, .25, .5, .75, 1.0),
                     labels = c('0%', '25%', '50%', '75%', '100%')) +
  geom_text(aes(label = sprintf('n = %s', n_tot_cat), y = 1.01), hjust = 0, 
            size = 2, 
            color = 'grey30') +
  coord_flip() +
  labs(x = 'Risk category', 
       y = 'Percent of species by quadrant', 
#       title = 'Risk category by quadrant',
       fill = 'Alignment')

>>>>

break_nums <- seq(0, 100, 20)

bar_taxa <- ggplot(taxa_df_long, aes(x = class, y = percent, fill = dataset)) +
  ggtheme_plot + 
  theme(panel.grid.major.x = element_blank()) +
  geom_bar(stat = 'identity', alpha = 1) +
  geom_hline(yintercept = break_nums, size = 0.25, color = 'white', alpha = .5) +
  scale_fill_manual(values = colvec) +
  geom_text(aes(label = sprintf('n = %s', n_total), y = 102), 
            size = 2, 
            angle = 90, hjust = 0, color = 'grey30') +
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 130),
                     breaks = break_nums,
                     labels = sprintf('%s%%', break_nums)) + 
  labs(# title = 'Species maps by taxonomic class',
       x = 'Class', 
       y = 'Percent of mapped species',
       fill = 'Dataset')

@@

``` {r plot bar-line charts by taxa, echo = FALSE, message = FALSE, warning = FALSE, eval = TRUE}
### bar charts comparing number of species in each class for each data set

taxa_df1 <- taxa_df %>%
  dplyr::select(class, AquaMaps = n_am_class_solo, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
  mutate(AquaMaps = ifelse(is.na(AquaMaps), 0, AquaMaps),
         IUCN     = ifelse(is.na(IUCN),     0, IUCN),
         Both     = ifelse(is.na(Both),     0, Both),
         n_total  = AquaMaps + IUCN + Both,
         IUCN_end = 100 * (IUCN + Both)/n_total,
         IUCN_start = 0,
         AquaMaps_end   = 100,
         AquaMaps_start = 100 * (n_total - AquaMaps - Both)/n_total,
         Both_end = AquaMaps_start,
         Both_start = IUCN_end) %>%
  gather(key = dataset,  value = start, ends_with('start')) %>%
  gather(key = dataset2, value = end, ends_with('end')) %>%
  mutate(dataset = str_replace(dataset, '_start', ''),
         dataset2 = str_replace(dataset2, '_end', '')) %>%
  filter(dataset == dataset2) %>%
  select(-dataset2)

taxa_all_df <- taxa_df1 %>%
  group_by(dataset) %>%
  summarize(class = 'Total',
            AquaMaps_total = sum(AquaMaps) + sum(Both),
            IUCN_total = sum(IUCN) + sum(Both),
            n_total  = sum(n_total),
            IUCN_end = 100 * IUCN_total/n_total,
            IUCN_start = 0,
            AquaMaps_end   = 100,
            AquaMaps_start = 100 * (n_total - AquaMaps_total)/n_total,
            Both_end = AquaMaps_start,
            Both_start = IUCN_end) %>%
  select(-dataset) %>% distinct() %>%
  gather(key = dataset,  value = start, ends_with('start')) %>%
  gather(key = dataset2, value = end, ends_with('end')) %>%
  mutate(dataset = str_replace(dataset, '_start', ''),
         dataset2 = str_replace(dataset2, '_end', '')) %>%
  filter(dataset == dataset2) %>%
  select(-dataset2, -AquaMaps_total, -IUCN_total)

taxa_df2 <- taxa_all_df %>%
  bind_rows(taxa_all_df %>%
              mutate(class = ' ',
                     start = 0,
                     end   = 0,
                     n_total = NA)) %>%
  bind_rows(taxa_df1 %>%
              select(-IUCN, -AquaMaps, -Both)) %>%
  mutate(class = factor(class, levels = unique(.$class), ordered = TRUE),
         n_label = ifelse(!is.na(n_total), paste0('n = ', n_total), ''))

break_nums <- seq(0, 100, 20)

dodge <- position_dodge(width = 0.1)

bar_taxa <- ggplot(taxa_df2, aes(x = class, color = dataset)) +
  ggtheme_plot + 
  theme(panel.grid.major.x = element_blank()) +
  geom_linerange(aes(ymin = start, ymax = end, size = dataset),
                 position = position_dodge(width = 0.5)) +
  scale_size_manual(values = c('IUCN' = .25, 'AquaMaps' = .25, 'Both' = 2.5)) +
  scale_color_manual(values = c('IUCN' = 'black', 'AquaMaps' = 'grey60', 'Both' = 'black')) +
  geom_text(aes(label = n_label, y = 102),
            size = 2,
            angle = 90, hjust = 0, color = 'grey30') +
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(0, 130),
                     breaks = break_nums,
                     labels = sprintf('%s%%', break_nums)) +
  labs(# title = 'Species maps by taxonomic class',
       x = 'Class',
       y = 'Percent of mapped species',
       fill = 'Dataset')

print(bar_taxa)
ggsave(filename = file.path(dir_fig, 'bar2_by_taxa_all.png'), 
       width = 17.8, height = 7, units = 'cm', dpi = 600)

![](../figures/bar2_by_taxa_all.png)

@@

break_nums <- seq(0, 100, 20)

### mongo plot time
scatter_quadplot <- ggplot(spp_list_labs, aes(x = x, y = y)) +
  ggtheme_plot + 
  ### color the quadrant backgrounds:
  annotate("rect", xmin = area_align_mean, xmax = 100, ymin = dist_align_mean, ymax = 100, 
           alpha = .3, fill= "#4dac26")  + 
  annotate("rect", xmax = area_align_mean, xmin =   0, ymin = dist_align_mean, ymax = 100, 
           alpha = .3, fill= "#b8e186") + 
  annotate("rect", xmin = area_align_mean, xmax = 100, ymax = dist_align_mean, ymin =   0, 
           alpha = .3, fill= "#f1b6da") + 
  annotate("rect", xmax = area_align_mean, xmin =   0, ymax = dist_align_mean, ymin =   0, 
           alpha = .3, fill= "#d01c8b") +
  ### add grid lines
  geom_vline(xintercept = break_nums, size = 0.25, color = 'white', alpha = .5) +
  geom_hline(yintercept = break_nums, size = 0.25, color = 'white', alpha = .5)

### plot the data points!
scatter_quadplot <- scatter_quadplot +
  # geom_point(color = '#4d4dac', aes(alpha = fade), show.legend = FALSE) +
  # geom_point(data = spp_list_labs %>%
  #              filter(str_detect(spp_group_text, 'Corals')),
  #            color = 'red', aes(alpha = fade), 
  #            shape = 4, stroke = .5, 
  #            size = 1, show.legend = FALSE) +
  # scale_alpha_manual(values = c('TRUE' = .15, 'FALSE' = 1)) +
  geom_point(color = 'black', aes(alpha = fade), size = .1, show.legend = FALSE) +
  scale_alpha_manual(values = c('TRUE' = .15, 'FALSE' = 1))

@@

### Manage axis scales 
scatter_quadplot <- scatter_quadplot +
  scale_x_continuous(expand = c(0, 0), limits = c(-1, 101),
                     breaks = break_nums,
                     labels = sprintf('%s%%', break_nums)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(-1, 101),
                     breaks = break_nums,
                     labels = sprintf('%s%%', break_nums))

@@

### Plot the bar chart
barchart_spp_gp_quads <- ggplot(spp_gp_quadrants %>% 
                                  transform(spp_group_text = reorder(spp_group_text, pct_q1)), 
                                aes(x = spp_group_text, fill = quad, weight = pct_quad)) +
  ggtheme_plot + 
  theme(panel.grid.major.y = element_blank()) +
  geom_bar(stat = 'count', alpha = 1) +
  scale_fill_manual(values = c('q4' = '#d01c8b', 
                               'q3' = '#f1b6da', 
                               'q2' = '#b8e186',
                               'q1' = '#4dac26'),
                    labels = quad_names$quad_name) +
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 1.1),
                     breaks = break_nums/100,
                     labels = sprintf('%s%%', break_nums)) + 
  ### add grid lines; horizontal but then get flipped
  geom_hline(yintercept = break_nums/100, size = 0.25, color = 'white', alpha = .5) +
  ### add text
  geom_text(# data = spp_gp_quadrants %>%
              # select(spp_group_text, n_spp) %>%
              # distinct(),
            aes(label = sprintf('n = %s', n_spp), y = 1.01), hjust = 0, 
            size = 2, 
            color = 'grey30') +
  coord_flip() +
  labs(x = 'Taxonomic Group', 
       y = 'Percent of species by quadrant', 
       fill = 'Alignment')

@@

### Plot the bar chart
barchart_spp_risk_quads <- ggplot(spp_risk_quadrants1 %>% 
                                  transform(popn_category = reorder(popn_category, cat_rank)), 
                                aes(x = popn_category, fill = quad, weight = pct_quad)) +
  ggtheme_plot + 
  theme(panel.grid.major.y = element_blank()) +
  geom_bar(stat = 'count', alpha = 1) +
  scale_fill_manual(values = c('q1 well-matched'   = '#4dac26', 
                               'q2 dist-matched'   = '#b8e186', 
                               'q3 area-matched'   = '#f1b6da', 
                               'q4 poorly matched' = '#d01c8b')) +
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 1.1),
                     breaks = break_nums/100,
                     labels = sprintf('%s%%', break_nums)) + 
  ### add grid lines
  geom_hline(yintercept = break_nums/100, size = 0.25, color = 'white', alpha = .5) +
  ### add text
  geom_text(aes(label = sprintf('n = %s', n_tot_cat), y = 1.01), hjust = 0, 
            size = 2, 
            color = 'grey30') +
  coord_flip() +
  labs(x = 'Risk category', 
       y = 'Percent of species by quadrant', 
#       title = 'Risk category by quadrant',
       fill = 'Alignment')

### 2016-08-10 ms_1_intro.Rmd
fixing up the plots: new quad plot, new coral quad plot, new dataset by taxa plot, new OHI plot

We then reexamine two recent marine biodiversity studies - an assessment of the status of global biodiversity within the Ocean Health Index [@halpern_index_2012; @halpern_patterns_2015] and a global analysis of gaps in protection afforded by marine protected areas (MPAs) [@klein_shortfalls_2015] - as case studies to explore the implications of prioritizing one data set over the other.  The results highlight possible consequences of different data use decisions on our understanding of marine biodiversity status and protection.

>>>>

We then reexamine two recent marine biodiversity studies - an assessment of the status of global biodiversity within the Ocean Health Index (OHI) [@halpern_index_2012; @halpern_patterns_2015] and a global analysis of gaps in protection afforded by marine protected areas (MPAs) [@klein_shortfalls_2015] - as case studies to explore the implications of prioritizing one data set over the other.  The results highlight possible consequences of different data use decisions on our understanding of marine biodiversity status and protection.

___we are not recommending one data set as better than the other!___

___acknowledge or debunk other global data sets - e.g. SAUP?___

### 2016-08-10 ms_2_methods.Rmd
fixing up the plots: new quad plot, new coral quad plot, new dataset by taxa plot, new OHI plot

Noting that a large proportion of corals were identified in the upper-left quadrant, we examined a handful of IUCN coral maps and found that each predicted coral presence in far deeper waters than would be expected based on the IUCN's habitat description.  To examine this issue further, we created a 200 m bathymetry raster from Natural Earth's 200 m bathymetry polygon [@natural_earth???] and masked our IUCN coral rasters to constrain the maps to shallow waters.  None of the paired-map corals is described as occurring deeper than 200 m, ___and most are found in far shallower waters.___  The resulting maps were again compared to the AquaMaps rasters; ___figure X___ plots shifts in coral species alignment due to this depth clipping.

**Methods for Ocean Health Index (OHI) case study:**
The Species component of the OHI Biodiversity goal (SPP) uses IUCN Red List extinction risk as a proxy for a species' health.  Each species is assigned a risk score based on IUCN Red List extinction risk category, from Least Concern (0.0) to Extinct (1.0) (Data Deficient and unassessed species are excluded).  Species range is determined from its IUCN range map, where available, or its AquaMaps range map ("presence" defined as a 40% or greater probability of occurrence) if an IUCN map is unavailable.  For each assessed OHI region, an area-weighted mean extinction risk across all species is calculated, and rescaled from 0 (mean extinction risk = 0.75) to 100 (mean extinction risk = 0) [@halpern_index_2012; @halpern_patterns_2015].

___For our analysis, we modified the original code for OHI 2015 Species status, allowing for flexibility in which data source was used for the overlapping species, and flexibility in modifying the AquaMaps presence threshold.  Using the output of the published OHI 2015 SPP status as our baseline, we ran the modified SPP code three times: once with IUCN  prioritizing IUCN over AquaMaps for a 0% threshold, and prioritizing AquaMaps over IUCN for both a 40% and 0% threshold.  We compared each of these to the output of the published OHI 2015 SPP status.___

_To highlight the differences in the datasets, we selected the subset of species whose ranges are represented in both datasets.  In the original OHI SPP analysis, IUCN data was used to determine ranges for these overlapping species; we calculated the SPP status for these species as our baseline.  We then recalculated using AquaMaps ranges at a 40% presence threshold (as used in the original OHI analysis) and a 0% threshold (to better simulate the "limits of distribution" defined by IUCN maps) and compared them to our IUCN baseline._

_We then modified the original OHI source code to recalculate the 2015 SPP status, using AquaMaps rather than IUCN for the overlapping species, and setting a 0% presence threshold on all AquaMaps species ranges.  This we compared to the original SPP status scores for 2015._

* ___overlapping species only? IUCN as baseline; AM 40%; AM 0%___

**Methods for MPA Gap Analysis case study:** 
To assess the effectiveness of MPAs in protecting biodiversity, Klein et al. [-@klein_shortfalls_2015] compared the coverage of the global MPA network presented by the World Database on Protected Areas (WDPA) [@iucn_world_2014] to the species ranges described in the 2014 AquaMaps dataset.  For this study, the researchers defined "presence" as 50% or greater probability of occurrence.

>>>>

Noting that a large proportion of corals were identified in the upper-left quadrant, we examined a handful of IUCN coral maps and found that each predicted coral presence in far deeper waters than would be expected based on the IUCN's habitat description.  Using the IUCN's API [REF] to find life history information for all IUCN-mapped corals, we found that none of the paired-map corals is described as occurring deeper than 200 m, ___and most are found in far shallower waters.___  We created a 200 m bathymetry raster from Natural Earth's 200 m bathymetry polygon [@natural_earth_XXX] and masked our IUCN coral rasters to exclude presence below 200 m.  The resulting maps were again compared to the AquaMaps rasters; ___figure XXXX___ plots shifts in coral species alignment due to this depth clipping. ___does some of this belong in the results/discussion?___

**Methods for Ocean Health Index (OHI) case study:**
The Species component of the OHI Biodiversity goal (SPP) uses IUCN Red List extinction risk as a proxy for a species' health.  Each species is assigned a risk score based on IUCN Red List extinction risk category, from Least Concern (0.0) to Extinct (1.0) (Data Deficient and unassessed species are excluded).  Species range is determined from its IUCN range map, where available, or its AquaMaps range map ("presence" defined as a 40% or greater probability of occurrence) if an IUCN map is unavailable.  For each assessed OHI region, an area-weighted mean extinction risk across all species is calculated, and rescaled to calculate a status score from 0 (mean extinction risk = 0.75) to 100 (mean extinction risk = 0) [@halpern_index_2012; @halpern_patterns_2015].

_To highlight the differences in the datasets, we selected the subset of species whose ranges are represented in both datasets, eliminating species found in only one or the other dataset.  In the original OHI SPP analysis, IUCN data was used to determine ranges for these overlapping species; we calculated all regional SPP status scores for just these species as our baseline.  We then recalculated the scores using AquaMaps ranges at a 40% presence threshold (as used in the original OHI analysis) and a 0% threshold (to better simulate the "limits of distribution" defined by IUCN maps) and compared them to our IUCN baseline._

_We then modified the original OHI source code to recalculate the 2015 SPP status scores for all regions and all species to examine the effects of changing the AquaMaps presence threshold from 40% to 0% (all range included), and the effects of changing the data source for the overlapping species from IUCN to AquaMaps.  We calculated the differences between these results and the published SPP status scores for 2015._

**Methods for MPA Gap Analysis case study:** 
To assess the effectiveness of MPAs in protecting biodiversity, Klein et al. [-@klein_shortfalls_2015] compared the coverage of the global MPA network presented by the World Database on Protected Areas (WDPA) [@iucn_world_2014] to the species ranges described in the 2014 AquaMaps dataset [@aquamaps_XXX].  For the primary analysis, the researchers defined species "presence" as 50% or greater probability of occurrence.
